---
title: "Reproducible Research (2)"
subtitle: Data import and manipulation
author: "Zhuanghua Shi"
date: "28 April 2016"
output: beamer_presentation
header-includes:
  - \usetheme{metropolis}
---
  
  ```{r setup, include=FALSE}
source('loadPackages.R')
knitr::opts_chunk$set(echo = FALSE)
options(xtable.comment=FALSE, fig.width = 4, fig.height=3, digits = 2) #hide  comments
```

## Github repository

> git clone https://github.com/strongway/seminar-reproducible-research.git

## Data import

- We encounter many types of data formats
    - csv, excel, matlab, spss, web table, Sql
- R {utils} provides basic import methods
    - read.csv
    - read.delim
    - read.table
- data.table package provides powerful function
    - fread()
- readr package (fast)
    - read_csv
    - read_delim
    - read_tsv
    
## import text file basic functions

> read.table(filename, header=TRUE, sep=',', stringsAsFactors = FALSE, col.names = c(...))

- specific functions:
    - read.csv()
    - read.csv2()
    - read.delim()
    - read.delim2()
- helper function
    - file.path()
    - past0()

## Comparison between utils and readr

utils          readr
------         ----------
read.table()   read_delim()
read.csv()     read_csv()
read.delim()   read_tsv()

## fread() from data.table

- super powerful, fast
- automatically manage column names
- infer column types and separators
- select or drop columns

## Practise 

Hands-on data from github

## Import matlab file

Two options for importing matlab data

- convert data format to csv within Matlab
- using package 'R.matlab' 
    - readMat

## Import multiple data

A good way to import multiple data set is:

- write a function of importing individual data
- using lapply() or Map() 

Some helper functions 

- dir() list the files in a specific folder
- do.call(rbind, a_list) combine data together

## Data cleaning and exploration

- data exploration
    - hist()
    - boxplot()
    - summary()

## Data manipulation

Two important packages

- dplyr - grammar-like manipulation
- data.table - sql-like manipulation

I prefer the former for its clearity.

## Functions in dplyr

- filter and arrange
- group_by
- select and mutate
- summarise
- pipe operator (Important feature)

## tidyr functions 

- gather(data, key, value, ...)
- spread(data, key, value)
- separate(data, key, colnames)
- unite(data, col, ..., sep)

## Course schedule
```{r, comment=NA,message=FALSE, results='asis'}
dates <- seq(as.Date('2016-04-14'), length.out = 6,by='2 weeks')
dates[6] <- dates[6] + 7 # move one week later due to conference
dates <- as.character(dates) # bug in xtable
contents <- c('Version control system - git', 
              'Data import and manipulation',
              'Publication-ready figures',
              'Statstics and Modeling',
              'Reusable data analysis',
              'RMarkdown and writing')
kable(data.frame(dates, contents))
```


